---
title: "team"
author: "Allegra Frickel"
date: "3/27/2019"
output: html_document
---

```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
```

```{r}
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)
```

```{r}
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(bio)
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

**Part One: Cleaning**

1. Player names and opponent names are factors
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)
offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)
bio$Name <- as.factor(bio$Name)
```

2. Offensive and defensive names statistics are numerical
```{r}
defense <- defense %>% mutate_if(is.character,as.numeric)
offense <- offense %>% mutate_if(is.character,as.numeric)
bio$Weight <- as.numeric(bio$Weight)
```

3. Change height to numeric

```{r}
bio <- bio %>% separate(Height, into = c("Feet", "Inches"), sep = "-")
```

```{r}
bio$Feet <- as.numeric(bio$Feet)
bio$Inches <- as.numeric(bio$Inches)
bio <- bio %>% mutate(HeightDec = Feet + Inches/12)
#would like to move HeightDec column after inches column
```

Print the structures of the Clean dataframes.
```{r}
defClean <- defense
offClean <- offense
bioClean <- bio
str(defClean)
str(offClean)
str(bioClean)
```

**Part Two: Tidying (1)**

1. Reformat defClean into a tidy format using gather. The type of statistic is added as a new key column names Statistic.

```{r}
defClean <- defClean %>% gather(key = Statistic, value = Count, 3:11)
```

2. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?

```{r}
ggplot(defClean, aes(x=Count)) + geom_histogram(binwidth = 1) + facet_wrap(~Statistic)
```

Based on the histrograms above, Turnover_FR (fumble recovery) and Turnover_INT (turnover interception) skills are rare.



3. Did ISU have better defense against West Virginia or KState? Create a scatterplot with the x-axis being the number of solo tackles against West Virginia. The y-axis should be against KState. A large number of solo tackles indicates a good defense.

```{r}
defClean %>% group_by(Opponent_Opponent) %>% filter(Opponent_Opponent == 'West Virginia' | Opponent_Opponent =='Kansas State')
plt <- ggplot(defClean, aes(x= Opponent_Opponent == `West Virginia`, y= Opponent_Opponent == `Kansas State`)) + geom_point() + xlab('Solo tackles against West Virgina') + ylab('Solo tackles against Kansas State')
#this line of code ran, but a scatterplort did not display. An error message saying West Virginia was not found.
```

Iowa State had a better defense against West Virginia/Kansas State. We can tell because Iowa State has higher plot points indicating a better defense against ___.

**Part Two: Tidying (2)**

1. Separate the Hometown column into two columns, name the city and the state. Print the first few rows of these two columns.

```{r}
bioClean <- bioClean %>% separate(Hometown, into = c("City", "State"), sep = ",")
head(bioClean)
```

2. How many players are there from each state?

```{r}
table(bioClean$State)
```

**Part Three: Joining data frames**

1. Find where the offensive players come from by joining the offensive data and player biography data appropriately.

```{r}
anti_join(bioClean, offClean, by = "Name")
```

2. Is there a relationship between the weight of the player and the receiving yards (Receiving_YDS)?

3. Is there a relationship between weight and solo tackles?


