---
title: "team"
author: "Allegra Frickel"
date: "3/27/2019"
output: html_document
---

```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2018.xlsx', sheet='Defensive')
str(defense)
```

```{r}
offense <- read_excel('cyclonesFootball2018.xlsx', sheet='Offensive')
str(offense)
```

```{r}
bio <- read_excel('cyclonesFootball2018.xlsx', sheet='Biography')
str(bio)
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

**Part One: Cleaning**

1. Player names and opponent names are factors
```{r}
defense$Name <- as.factor(defense$Name)
defense$Opponent_Opponent <- as.factor(defense$Opponent_Opponent)
offense$Name <- as.factor(offense$Name)
offense$Opponent_Opponent <- as.factor(offense$Opponent_Opponent)
bio$Name <- as.factor(bio$Name)
```

2. Offensive and defensive names statistics are numerical
```{r}
defense <- defense %>% mutate_if(is.character,as.numeric)
offense <- offense %>% mutate_if(is.character,as.numeric)
bio$Weight <- as.numeric(bio$Weight)
```

3. Change height to numeric
```{r}

```

Print the structures of the Clean dataframes.
```{r}
defClean <- defense
offClean <- offense
bioClean <- bio
str(defClean)
str(offClean)
str(bioClean)
```


##**Part Two: Tidying (1)**

1. Reformat defClean into a tidy format using gather. The type of statistic is added as a new key column names Statistic.

```{r}
library(tidyverse)
defClean %>% gather(key = Statistic, value = Count, 3:11)
```

2. Create a histogram for each defensive statistic. Use facet_wrap to draw all histograms in a single plot. What defensive skills are rare?
```{r}
#this is completely and totally incorrect
ggplot(gather(defClean), aes(x=value)) + 
    geom_bar() + 
    facet_wrap(~key)
```


3. Did ISU have better defense against West Virginia or KState? Create a scatterplot with the x-axis being the number of solo tackles against West Virginia. The y-axis should be against KState. A large number of solo tackles indicates a good defense.

(What Audrey tried)
```{r}
var1 <- defClean %>% group_by(Name) %>%
  filter(Opponent_Opponent == 'West Virginia' | Opponent_Opponent =='Kansas State') %>%
  select(Name, Opponent_Opponent, Tackles_Solo)

#ggplot(var1, aes(x=Opponent_Opponent, y=Tackles_Solo)) +geom_point()

```

(what Ally tried)
```{r}
plt <- ggplot(defClean, aes(x= 'West Virginia', y='Kansas State')) + geom_point() + xlab('Solo tackles against West Virgina') + ylab('Solo tackles against Kansas State')
plt
```

Iowa State had a better defense against West Virginia/Kansas State. We can tell because ____ has a large number of solo tackles of ___, which was during the West Virgina/Kansas State game.


**Part Two: Tidying (2)**

1. Separate the Hometown column into two columns, name the city and the state. Print the first few rows of these two columns.

```{r}
bioClean %>% separate(Hometown, into = c("City", "State"))
```

2. How many players are there from each state?

```{r}
table(bioClean$State)
```

**Part Three: Joining data frames**

1. Find where the offensive players come from by joining the offensive data and player biography data appropriately.

```{r}
anti_join(bioClean, offClean, by = "Name")
```

2. Is there a relationship between the weight of the player and the receiving yards (Receiving_YDS)?

3. Is there a relationship between weight and solo tackles?


